---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("G:/My Drive/academics/2122_uic_ms/2022_spring/ids506_health/prostate")

library(tidyverse)
library(tidymodels)
library(patchwork)
library(lubridate)
library(tidytext)
library(themis)

dat <- read_csv("data/training_data.csv") %>% 
  rename(y = survival_7_years)

dat <- dat %>% 
  filter(survival_1_year == 1)
```


```{r}
na_table <- function(d){
  NAs <- colMeans(is.na(d)) %>% data.frame()
  colnames(NAs) <- c("prop_NA")
  NAs <- NAs %>% arrange(desc(prop_NA))
  
  return(NAs)
}

na_table(dat)
```

# Functions

```{r}
crosstab <- function(dframe, column, y_col) {
  dframe %>%
    group_by(!!sym(column), !!sym(y_col)) %>%
    summarise(n = n()) %>%
    pivot_wider(id_cols = !!sym(column),
                names_from = !!sym(y_col),
                values_from = n) %>% 
    mutate(
      `0` = ifelse(is.na(`0`)==1,0,`0`),
      `1` = ifelse(is.na(`1`)==1,0,`1`),
    )
}

xtab <- function(dframe, column, y_col){
  xtab <- crosstab(dframe, column, y_col) %>% 
    mutate(survival_ratio = `1` / (`1` + `0`))
  return(xtab)
}

distro_table <- function(dframe, column, y_col) {
  xtab <- crosstab(dframe, column, y_col)
  
  plot0 <- xtab %>%
    pivot_longer(cols = -!!sym(column),
                 names_to = "yvar",
                 values_to = "val") %>%
    ggplot() +
    aes(x = round(!!sym(column)),
        y = val,
        col = yvar %>% factor(),
        fill = yvar %>% factor()) +
    scale_fill_viridis_d() + scale_color_viridis_d() +
    theme_minimal() 
  
  p1 <- plot0 + geom_col(position = "stack")+
    labs(col = "Target Variable",
         fill = "Target Variable",
         y = "Count") +
    theme(legend.position = "none")
  p2 <- plot0 + geom_col(position = "fill")+
    labs(col = "Target Variable",
         fill = "Target Variable",
         y = "Proportion")
  p <- (p1 | p2)
  return(p)
}

var_na <- function(dframe, column, y_col) {
  xtab <- crosstab(dframe, column, y_col)
  newcolname <- paste(column, sep = "")
  
  out <- xtab %>%
    mutate(is_na = ifelse(is.na(!!sym(column)) == 1, 1, 0)) %>%
    group_by(is_na) %>%
    summarise(`0` = sum(`0`),
              `1` = sum(`1`)) %>%
    mutate(n = (`1` + `0`),
           survival_ratio = `1` / (`1` + `0`))
  
  colnames(out)[1] <- newcolname
  
  return(out)
}
```

```{r}
get_na_table <- function(dframe, column, y_col){
  dt <- var_na(dframe, column, y_col) %>% select(!!sym(column),n,survival_ratio)
  
  col <- c(col = column,
    dt[2,2]/(sum(dt[2])),
    dt[1,3],
    dt[2,3]) %>% 
  as.data.frame() %>% 
  rename(NAs = n,
         rate.rest = survival_ratio,
         rate.NAs  = survival_ratio.1)
  
  return(col)
}

all_na_table <- function(dframe,y_col){
  names <- colnames(dframe)
  
  table <- data.frame()
  
  for (i in names){
    if (i == y_col){
      break
    } else {
      table <- table %>% rbind(
        get_na_table(dframe,i,y_col)
      )
    }
  }
  
  return(table)
}

all_crosstabs <- function(dframe,y_col){
  names <- colnames(dframe)
  table <- data.frame()
  
  for (i in names){
    if (i == y_col){
      next
    } else {
      newtab <- crosstab(dframe,i,y_col) %>% 
        mutate(
          variable = i
        ) %>% 
        rename(
          value = !!sym(i)
        ) %>% 
        mutate(
          value = value %>% factor()
        ) %>% 
        select(variable,everything())
      
      table <- table %>% rbind(newtab)
    }
  }
  return(table)
}
```

```{r}
dist_plot_d <- function(dframe,col,y){
  plot0 <- dframe %>% 
    group_by(!!sym(col),!!sym(y)) %>% 
    summarise(n=n()) %>% 
    ggplot()+
    aes(
      x = !!sym(col) %>% factor(),
      y = n,
      fill = !!sym(y) %>% factor(),
      col  = !!sym(y) %>% factor()
    )+
    scale_fill_viridis_d() + scale_color_viridis_d() +
    theme_minimal()
  
  p1 <- plot0 + geom_col(position = "stack")+
    labs(col = "Target Variable",
         fill = "Target Variable",
         y = "Count")+
    theme(legend.position = "none")
  p2 <- plot0 + geom_col(position = "fill")+
    labs(col = "Target Variable",
         fill = "Target Variable",
         y = "Proportion")
  p <- (p1 | p2)
  return(p)
}

dist_plot_c <- function(dframe,col,y){
  plot0 <- dframe %>% 
    group_by(!!sym(col),!!sym(y)) %>% 
    summarise(n=n()) %>% 
    ggplot()+
    aes(
      x = !!sym(col) %>% as.numeric(),
      y = n,
      fill = !!sym(y) %>% factor(),
      col  = !!sym(y) %>% factor()
    )+
    scale_fill_viridis_d() + scale_color_viridis_d() +
    theme_minimal()
  
  p1 <- plot0 + geom_col(position = "stack")+
    labs(col = "Target Variable",
         fill = "Target Variable",
         y = "Count")+
    theme(legend.position = "none")
  p2 <- plot0 + geom_col(position = "fill")+
    labs(col = "Target Variable",
         fill = "Target Variable",
         y = "Proportion")
  p <- (p1 | p2)
  return(p)
}
```



```{r eval=FALSE, include=T}
all_na_tables <- all_na_table(dat,"y")
all_na_tables[is.na(all_na_tables)] <- 0 
all_na_tables

write_csv(all_na_tables, "data/all_na_tables.csv")
```
# tumor_6_months

```{r}
xtab(dat,"tumor_6_months","y")
var_na(dat,"tumor_6_months","y")
distro_table(dat,"tumor_6_months","y")
```

# psa_6_months

```{r}
xtab(dat,"psa_6_months","y")
var_na(dat,"psa_6_months","y")
distro_table(dat,"psa_6_months","y")
```

# psa_1_year

```{r}
xtab(dat,"psa_1_year","y")
var_na(dat,"psa_1_year","y")
distro_table(dat,"psa_1_year","y")
```

# tumor_1_year

```{r}
xtab(dat,"tumor_1_year","y")
var_na(dat,"tumor_1_year","y")
distro_table(dat,"tumor_1_year","y")
```

# family_history

```{r}
xtab(dat,"family_history","y")
var_na(dat,"family_history","y")
distro_table(dat,"family_history","y")
```

# first_degree_history

```{r}
xtab(dat,"first_degree_history","y")
var_na(dat,"first_degree_history","y")
distro_table(dat,"first_degree_history","y")
```

# previous_cancer

```{r}
xtab(dat,"previous_cancer","y")
var_na(dat,"previous_cancer","y")
distro_table(dat,"previous_cancer","y")
```

# smoker

```{r}
xtab(dat,"smoker","y")
var_na(dat,"smoker","y")
distro_table(dat,"smoker","y")
```

# tea

```{r}
xtab(dat,"tea","y")
var_na(dat,"tea","y")
distro_table(dat,"tea","y")
```

# psa_diagnosis

```{r}
xtab(dat,"psa_diagnosis","y")
var_na(dat,"psa_diagnosis","y")
distro_table(dat,"psa_diagnosis","y")
```
# height

```{r}
xtab(dat,"height","y")
var_na(dat,"height","y")
distro_table(dat,"height","y")
```

# weight

```{r}
xtab(dat,"weight","y")
var_na(dat,"weight","y")
distro_table(dat,"weight","y")
```

# age

```{r}
xtab(dat,"age","y")
var_na(dat,"age","y")
distro_table(dat,"age","y")
```

# gleason_score

```{r}
xtab(dat,"gleason_score","y")
var_na(dat,"gleason_score","y")
distro_table(dat,"gleason_score","y")
```

# tumor_diagnosis

```{r}
xtab(dat,"tumor_diagnosis","y")
var_na(dat,"tumor_diagnosis","y")
distro_table(dat,"tumor_diagnosis","y")
```

# race

```{r}
xtab(dat,"race","y")
var_na(dat,"race","y")
distro_table(dat,"race","y")
```
# Removing NAs

```{r}
dat_no_na <- dat %>% 
  filter(
    !is.na(weight),
    !is.na(height),
    !is.na(age),
    !is.na(gleason_score),
    !is.na(tumor_diagnosis),
    !is.na(race)
  ) %>% 
  mutate(
    psa_1_year           = ifelse(is.na(psa_1_year)==1,0,psa_1_year),
    tumor_1_year         = ifelse(is.na(tumor_1_year)==1,0,tumor_1_year),
    family_history       = ifelse(is.na(family_history)==1,0,family_history),
    first_degree_history = ifelse(is.na(first_degree_history)==1,0,first_degree_history),
    previous_cancer      = ifelse(is.na(previous_cancer)==1,0,previous_cancer),
    smoker               = ifelse(is.na(smoker)==1,0,smoker),
    tea                  = ifelse(is.na(tea)==1,0,tea),
    psa_diagnosis        = ifelse(is.na(psa_diagnosis)==1,0,psa_diagnosis)
  ) %>% 
  select(-c(id,tumor_6_months,psa_6_months))
```

# Dealing with Dates

```{r}
dat_no_na %>% 
  transmute(
    zero_day = ifelse(str_detect(diagnosis_date,"00"),1,0),
    diagnosis_date = ifelse(zero_day == 1, diagnosis_date, NA),
    y = y
  ) %>% 
  group_by(zero_day,y) %>% 
  summarise(n=n()) %>% 
  mutate(sum = sum(n),
         prop = n/sum)
  
1117/(1117+10442)
```

```{r}
dat_no_na <- dat_no_na %>% 
  mutate(
    # Replacing 00 date values with 01
    zero_day = ifelse(str_detect(diagnosis_date,"00"),1,0),
    diagnosis_date = ifelse(zero_day == 0,
                            diagnosis_date,
                            str_replace(diagnosis_date, "00","01")),
    diagnosis_date = paste(diagnosis_date,"2021",sep="") %>% mdy() %>% yday()
  ) %>% 
  select(-zero_day) 
```

# discretes

```{r}
dat_no_na_d <- dat_no_na %>% 
  select(t_score,
         n_score,
         m_score,
         stage,
         race,
         family_history,
         first_degree_history,
         previous_cancer,
         smoker,
         side,
         tea,
         rd_thrpy,
         h_thrpy,
         chm_thrpy,
         cry_thrpy,
         brch_thrpy,
         rad_rem,
         multi_thrpy,
         survival_1_year,
         y)

all_crosstabs(dat_no_na_d,"y") %>% 
  write_csv("data/discrete_tabs.csv")
```


# diagnosis_date

```{r}
dist_plot_d(dat_no_na,"diagnosis_date","y")
```

# gleason_score

```{r}
dist_plot_c(dat_no_na,"gleason_score","y")
```

# age

```{r}
dist_plot_c(dat_no_na,"age","y")
```

# height

```{r}
dist_plot_c(dat_no_na,"height","y")
```

# weight

```{r}
dist_plot_c(dat_no_na,"weight","y")
```

# tumor_diagnosis

```{r}
dist_plot_c(dat_no_na,"tumor_diagnosis","y")
```

# tumor_6_months

```{r}
dist_plot_c(dat,"tumor_6_months","y")
```

# tumor_1_year

```{r}
dist_plot_c(dat_no_na,"tumor_1_year","y")
```

# psa_diagnosis

```{r}
dist_plot_c(dat_no_na,"psa_diagnosis","y")
```

# psa_6_months

```{r}
dist_plot_c(dat,"psa_6_months","y")
```

# psa_1_year

```{r}
dist_plot_c(dat_no_na,"psa_1_year","y")
```

# symptoms

```{r}
symptom_string <- paste(dat_no_na$symptoms,collapse = ',') %>% 
  str_squish() %>% 
  str_replace(" ","") %>% 
  strsplit(",") %>% .[[1]] %>% 
  matrix() %>% 
  as.data.frame() %>% 
  group_by(V1) %>% summarise() %>% 
  filter(V1 != "NA") %>% rename(symptom = V1)
```

```{r}
dat_no_na_symptom <- dat_no_na

stringcheck <- function(dframe,str){
  newcolname = paste0("symptom_",str,sep="")
  dframe <- dframe %>% 
    transmute(
      !!sym(newcolname) := ifelse(str_detect(symptoms,str),1,0)
    )
  return(dframe)
}

for (i in (1:nrow(symptom_string[1]))){
  str = paste(symptom_string[i,1])
  dat_no_na_symptom <- dat_no_na_symptom %>% 
    cbind(
      stringcheck(dat_no_na_symptom,str = str)
    )
}

dat_no_na_symptom <- dat_no_na_symptom %>% 
  select(-symptoms)

dat_no_na_symptom[is.na(dat_no_na_symptom)==1] <- 0
```

```{r}
library(textrecipes)

recipe(y~.,data = dat_no_na) %>% 
  step_tokenize(symptoms,options = list(strip_punct = T,
                                        lowercase = T)) %>% 
  step_tf(symptoms) %>% 
  prep() %>% 
  bake(dat_no_na)
```

# Cormap

```{r}
dat_no_na_symptom_c <- dat_no_na %>%
  select(-c(t_score,n_score,m_score,stage,race,family_history,first_degree_history,
            previous_cancer,smoker,side,tea,rd_thrpy,h_thrpy,chm_thrpy,cry_thrpy,brch_thrpy,
            rad_rem,multi_thrpy,survival_1_year,symptoms))

library(ggcorrplot)



dat_no_na_symptom_c %>% 
  cor() %>% 
  round(digits = 4) %>% 
  ggcorrplot(method = "circle",
             type = "upper",
             p.mat = dat_no_na_symptom_c %>% cor_pmat(),
             ggtheme = ggplot2::theme_minimal,
             colors = c("#fde725","#ffffff","#440154"),
             hc.order = FALSE)
```

# Symptom Tabs

```{r}
dat_symptoms_only <- dat_no_na_symptom %>% 
  select(
    y,
    starts_with("symptom")
  )

all_crosstabs(dat_symptoms_only,"y") %>% 
  write_csv("data/symptom_tabs.csv")
```
