---
title: "Regression Modeling"
output: html_document
---

<style type="text/css">
.main-container {
  max-width: 1920px;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Import packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(gt)
library(ggridges)
library(lm.beta)

options(scipen=999)
```

# Import data 

```{r message=FALSE, warning=FALSE}
# Data import (will load selector widget to choose local file)
xlsx_import <- read_excel(file.choose(),sheet = "data")
```

# Cleaning data

```{r message=FALSE, warning=FALSE}
# Data cleaning step
dat <- xlsx_import %>% 
  select(-c(url,url_id,transcript)) %>% 
  clean_names() %>% 
  mutate(
    
    # Parsing video length string as number of seconds
    vlength_s = str_match(video_length,"\\d+$") %>% as.integer(),
    vlength_m = str_match(str_remove(video_length,":\\d+$"),"\\d+$") %>% as.integer(),
    vlength_h = str_remove(video_length,"\\d+:\\d+$") %>% str_remove(":") %>% as.integer(),
    vlength_h = ifelse(is.na(vlength_h),0,vlength_h),
    video_length = (vlength_s + vlength_m*60 + vlength_h*3600)/60,
    
    # Parsing missing values in sub count, negative comment count, and neutral comment count as 0
    subs_count = subs_count %>% as.integer(),
    subs_count = ifelse(is.na(subs_count),0,subs_count),
    
    cmt_negative_n = cmt_negative_n %>% as.integer(),
    cmt_negative_n = ifelse(is.na(cmt_negative_n),0,cmt_negative_n),
    
    cmt_neutral_n = cmt_neutral_n %>% as.integer(),
    cmt_neutral_n = ifelse(is.na(cmt_neutral_n),0,cmt_neutral_n)
    ) %>% 
  select(-c(vlength_s,vlength_m,vlength_h)) 

# parsing transcript emotion data to most prominent emotion
text_emotions <- dat %>% 
  select(video_id,starts_with("text")) %>% 
  pivot_longer(cols = starts_with("text"),
               names_to = "names",
               values_to = "val") %>% 
  group_by(video_id) %>% 
  filter(val == max(val)) %>% 
  transmute(
    video_id = video_id,
    prom_text_emote = str_replace(names,"text_"," ")
  )

dat <- dat %>% left_join(text_emotions)
```

# Plotting

```{r}
dat %>% 
  filter(gender != 2) %>% 
  mutate(subs_count = (subs_count+1)) %>% 
  ggplot()+
  geom_density_ridges(aes(x = subs_count, 
                          y = factor(gender),
                          fill = factor(gender),
                          col = factor(gender)),
                      alpha = .5)+
  theme_minimal()+
  labs(x = "Subscriber Count (log scale)",
       y = "Density",
       col = "Gender",
       fill = "Gender")+
  scale_x_log10()+
  scale_fill_viridis_d()+
  scale_color_viridis_d()
```

```{r}
draw_density_plot_c <- function(dframe,var,log=0){
  if(log == 0){
    plot <- dframe %>% 
      mutate(xvar = !!sym(var)) %>% 
      ggplot()+
      aes(x = xvar)+
      geom_histogram(fill = "#21918c",
                     col  = "#21918c",
                     alpha = .75,
                     bins = 30)+
      theme_minimal()+
      labs(x = paste(var,"density (20 bins)"),
           y = "Count")
  }else if(log == 1){
    plot <- dframe %>% 
      mutate(xvar = !!sym(var) + 1) %>% 
      ggplot()+
      aes(x = xvar)+
      geom_histogram(fill = "#21918c",
                     col  = "#21918c",
                     alpha = .75,
                     bins = 30)+
      theme_minimal()+
      scale_x_log10()+
      labs(x = paste(var,"density (logscale + 20 bins)"),
           y = "Count")
  }
  
  return(plot)
}


cols_d <- dat %>% select(-c(video_id,
                         covid_period,
                         gender,
                         people_in_video)) %>% 
  select(-starts_with(c("prom","topic","emote"))) %>% 
  colnames()

```

```{r}
draw_density_plot_c(dat,"video_length")
draw_density_plot_c(dat,"subs_count",1)
draw_density_plot_c(dat,"n_likes",1)
draw_density_plot_c(dat,"n_views",1)
draw_density_plot_c(dat,"cmt_n",1)
draw_density_plot_c(dat,"cmt_wordcount",1)
draw_density_plot_c(dat,"cmt_nltk_sentiment")

# These are fluff
#draw_density_plot_c(dat,"cmt_positive_n")
#draw_density_plot_c(dat,"cmt_positive_mean")
#draw_density_plot_c(dat,"cmt_positive_med")
#draw_density_plot_c(dat,"cmt_negative_n")
#draw_density_plot_c(dat,"cmt_negative_mean")
#draw_density_plot_c(dat,"cmt_negative_med")
#draw_density_plot_c(dat,"cmt_neutral_n")
#draw_density_plot_c(dat,"cmt_neutral_mean")
#draw_density_plot_c(dat,"cmt_neutral_med")

```


# Modeling function

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Function to get coef + sig table for an x df, a y df, and a selected variable string
f.linreg <- function(dfx,dfy,yvar){
  
  # Joining x and y frames
  df_temp <-
    dfy %>% 
    select(video_id, y = !!sym(yvar)) %>% 
    left_join(dfx, by = 'video_id') %>% 
    select(-video_id)
  
  # Fit model
  model_temp <- lm(y~.,
                   data = df_temp#, singular.ok = FALSE
                   )
  
  # Extract coefficients
  coefs_temp <- model_temp$coefficients %>% as.data.frame()
  pvals_temp <- summary(model_temp)$coefficients[,4] %>% as.data.frame()
  
  
  colnames(coefs_temp) <- paste('coef')
  colnames(pvals_temp) <- "p"
    
  # Converting p-values to asterisk significance notation
  pvals_temp <- pvals_temp %>% mutate(p = case_when(p <= .001 ~ "*",
                                                    p <= .01  ~ "**",
                                                    p <= .05  ~ "***",
                                                    p > .05   ~ " "))

  # Binding coefficients and stat.sig frames
  output <- coefs_temp %>% rownames_to_column("Feature") %>% 
    left_join(pvals_temp %>% rownames_to_column("Feature")) %>% 
    transmute(Feature,
              coef = paste(round(coef,digits=3),p))
  
  colnames(output) <- c("Feature",yvar)
  return(output)
}

f.linreg.s <- function(dfx,dfy,yvar){
  
  # Joining x and y frames
  df_temp <-
    dfy %>% 
    select(video_id, y = !!sym(yvar)) %>% 
    left_join(dfx, by = 'video_id') %>% 
    select(-video_id)
  
  # Fit model
  model_temp <- lm(y~.,
                   data = df_temp#, singular.ok = FALSE
                   )
  
  b = lm.beta(model_temp)
  
  coefs_temp <- b$standardized.coefficients %>% as.data.frame()
  pvals_temp <- summary(model_temp)$coefficients[,4] %>% as.data.frame()
  
  
  colnames(coefs_temp) <- paste('coef')
  colnames(pvals_temp) <- "p"
    
  # Converting p-values to asterisk significance notation
  pvals_temp <- pvals_temp %>% mutate(p = case_when(p <= .001 ~ "*",
                                                    p <= .01  ~ "**",
                                                    p <= .05  ~ "***",
                                                    p > .05   ~ " "))

  # Binding coefficients and stat.sig frames
  output <- coefs_temp %>% rownames_to_column("Feature") %>% 
    left_join(pvals_temp %>% rownames_to_column("Feature")) %>% 
    transmute(Feature,
              coef = paste(round(coef,digits=3),p))
  
  colnames(output) <- c("Feature",yvar)
  return(output)
}
```

# Without FER data
## Separating into Features and Target Variables

```{r message=FALSE, warning=FALSE}
# Removing FER emotion scores
dat_no_emote <- dat %>% 
  select(-starts_with("emote")) %>% 
  select(-c(prom_emote)) %>% 
  select(-starts_with("topic"))

# Separating into one df for target vars and one for feature vars
df_y <- dat_no_emote %>% select(video_id,n_likes,n_views,starts_with("cmt_"),starts_with('proportion'))
df_x <- dat_no_emote %>% select(-c(n_likes,n_views),-starts_with("cmt_"),-starts_with('proportion'),-starts_with("text"))
```

## Applying modeling function to Target Variables

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
y_names <- colnames(df_y) %>% as.vector()
y_names = y_names[y_names!='video_id']

for(i in 1:length(y_names)){
  var = y_names[i]
  if(i == 1){
    df0 = f.linreg(df_x,df_y,var)
  }else{
    dfn = f.linreg(df_x,df_y,var)
    df0 <- left_join(df0,dfn,by = 'Feature')
    rm(dfn)
  }
}

for(i in 1:length(y_names)){
  var = y_names[i]
  if(i == 1){
    df0_standard = f.linreg.s(df_x,df_y,var)
  }else{
    dfn = f.linreg.s(df_x,df_y,var)
    df0_standard <- left_join(df0_standard,dfn,by = 'Feature')
    rm(dfn)
  }
}

```


```{r message=FALSE, warning=FALSE, paged.print=TRUE}
builder <- function(x, Limit){cells_body(columns = !!sym(x), rows = str_detect(!!sym(x),"\\*"))}

gt0 <- df0 %>% 
  # select(!matches("_mean")) %>%  
  # select(!matches("med")) %>% 
  mutate(across(where(is.numeric), ~ round(., 3))) %>% 
  gt(rowname_col = "Feature") %>% 
  tab_header(title = "Unstandardized Regression Coefficients for all Target Variables",
             subtitle = "without Facial Emotion Recognition (FER)") %>% 
  tab_stubhead(label = "Feature") %>% 
  tab_style(
     locations = cells_column_labels(columns = everything()),
     style     = list(
       #Give a thick border below
       cell_borders(sides = "bottom", weight = px(2)),
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  tab_style(
     locations = lapply(colnames(df0 %>% select(-Feature)),builder),
     style     = list(
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  #Apply different style to the title
  tab_style(
    locations = cells_title(groups = "title"),
    style     = list(cell_text(weight = "bold", size = 24))
  ) %>% 
  opt_all_caps() %>% 
  opt_table_font(
    font = list(google_font("Raleway"),
                default_fonts())
  )

gt0_standard <- df0_standard %>% 
  # select(!matches("_mean")) %>%  
  # select(!matches("med")) %>% 
  mutate(across(where(is.numeric), ~ round(., 3))) %>% 
  gt(rowname_col = "Feature") %>% 
  tab_header(title = "Standardized Regression Coefficients for all Target Variables",
             subtitle = "without Facial Emotion Recognition (FER)") %>% 
  tab_stubhead(label = "Feature") %>% 
  tab_style(
     locations = cells_column_labels(columns = everything()),
     style     = list(
       #Give a thick border below
       cell_borders(sides = "bottom", weight = px(2)),
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  tab_style(
     locations = lapply(colnames(df0_standard %>% select(-Feature)),builder),
     style     = list(
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  #Apply different style to the title
  tab_style(
    locations = cells_title(groups = "title"),
    style     = list(cell_text(weight = "bold", size = 24))
  ) %>% 
  opt_all_caps() %>% 
  opt_table_font(
    font = list(google_font("Raleway"),
                default_fonts())
  )
```

## Printing + saving summary tables
```{r}
gtsave(gt0,"regression_outputs_no_FER.html")
gtsave(gt0_standard,"regression_outputs_no_FER_standardized.html")

gt0
gt0_standard
```

# With FER data
## Separating into Features and Target Variables

```{r message=FALSE, warning=FALSE}
# Removing FER emotion scores
dat_w_emote <- dat %>% 
  select(-starts_with("emote")) %>% 
  select(-starts_with("topic")) %>% 
  select(-people_in_video) %>% 
  mutate(prom_emote = str_replace(prom_emote,"^emote_","")) %>% 
  filter(!is.na(prom_emote))

# Separating into one df for target vars and one for feature vars
df_y_em <- dat_w_emote %>% select(video_id,n_likes,n_views,starts_with("cmt_"),starts_with("proportion"))
df_x_em <- dat_w_emote %>% select(-c(n_likes,n_views),-starts_with("cmt_"),-starts_with("proportion"),-starts_with('text'))
```

```{r}
intersect(colnames(df_y_em),
          colnames(df_x_em))
```


## Applying modeling function to Target Variables

```{r}
y_names_em <- colnames(df_y_em) %>% as.vector()
y_names_em = y_names_em[y_names_em!='video_id']

for(i in 1:length(y_names_em)){
  var = y_names_em[i]
  if(i == 1){
    df1 = f.linreg(df_x_em,df_y_em,var)
  }else{
    dfn = f.linreg(df_x_em,df_y_em,var)
    df1 <- left_join(df1,dfn,by = 'Feature')
    rm(dfn)
  }
}

for(i in 1:length(y_names_em)){
  var = y_names_em[i]
  if(i == 1){
    df1_standard = f.linreg.s(df_x_em,df_y_em,var)
  }else{
    dfn = f.linreg.s(df_x_em,df_y_em,var)
    df1_standard <- left_join(df1_standard,dfn,by = 'Feature')
    rm(dfn)
  }
}


gt1 <- df1 %>% 
  # select(!matches("_mean")) %>%  
  # select(!matches("med")) %>% 
  mutate(across(where(is.numeric), ~ round(., 3))) %>% 
  gt(rowname_col = "Feature") %>% 
  tab_header(title = "Unstandardized Regression Coefficients for all Target Variables",
             subtitle = "with Facial Emotion Recognition (FER)") %>% 
  tab_stubhead(label = "Feature") %>% 
  tab_style(
     locations = cells_column_labels(columns = everything()),
     style     = list(
       #Give a thick border below
       cell_borders(sides = "bottom", weight = px(2)),
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  tab_style(
     locations = lapply(colnames(df1 %>% select(-Feature)),builder),
     style     = list(
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  #Apply different style to the title
  tab_style(
    locations = cells_title(groups = "title"),
    style     = list(cell_text(weight = "bold", size = 24))
  ) %>% 
  opt_all_caps() %>% 
  opt_table_font(
    font = list(google_font("Raleway"),
                default_fonts())
  )

gt1_standard <- df1_standard %>% 
  # select(!matches("_mean")) %>%  
  # select(!matches("med")) %>% 
  mutate(across(where(is.numeric), ~ round(., 3))) %>% 
  gt(rowname_col = "Feature") %>% 
  tab_header(title = "Standardized Regression Coefficients for all Target Variables",
             subtitle = "with Facial Emotion Recognition (FER)") %>% 
  tab_stubhead(label = "Feature") %>% 
  tab_style(
     locations = cells_column_labels(columns = everything()),
     style     = list(
       #Give a thick border below
       cell_borders(sides = "bottom", weight = px(2)),
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  tab_style(
     locations = lapply(colnames(df1_standard %>% select(-Feature)),builder),
     style     = list(
       #Make text bold
       cell_text(weight = "bold")
     )
  ) %>% 
  #Apply different style to the title
  tab_style(
    locations = cells_title(groups = "title"),
    style     = list(cell_text(weight = "bold", size = 24))
  ) %>% 
  opt_all_caps() %>% 
  opt_table_font(
    font = list(google_font("Raleway"),
                default_fonts())
  )

```

## Printing + saving summary tables

```{r}
gtsave(gt1,"regression_outputs_w_FER.html")
gtsave(gt1_standard,"regression_outputs_w_FER_standardized.html")

gt1
gt1_standard
```