---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Env

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(tidymodels)
library(themis)

setwd("G:/My Drive/academics/2122_uic_ms/2022_spring/ids506_health/kidney")
```

# Import

```{r}
dat <- read_csv("data.csv") %>% 
  filter(is.na(CKD) == FALSE)

NAs <- colMeans(is.na(dat)) %>% data.frame()
colnames(NAs) <- c("prop_NA")
NAs %>% 
  arrange(desc(prop_NA))

df <- dat %>% 
  filter_all(all_vars(is.na(.)==F))

df <- df %>% 
  select(
    # Output Var
    CKD,
    
    # Demo
    Age,
    Female,
    Racegrp,
    Educ,
    Unmarried,
    Income,
    
    # Healthcare Access
    CareSource,
    Insured,
    
    # Metrics (eliminating weight + height as BMI reduces both)
    # (eliminating TotalChol = HDL+LDL as granular values are being kept)
    BMI,
    Obese,
    Waist,
    SBP,
    DBP,
    HDL,
    LDL,
    Dyslipidemia,
    PVD,
    
    # Lifestyle & Family History
    Activity,
    PoorVision,
    Smoker,
    Hypertension,
    Hypertension_fam = `Fam Hypertension`,
    Diabetes,
    Diabetes_fam = `Fam Diabetes`,
    Stroke,
    CVD,
    CVD_fam = `Fam CVD`,
    CHF,
    Anemia
  ) %>% 
  mutate(
    CKD = factor(CKD),
    Activity = factor(Activity),
    
  )
```

```{r}
write_csv(df,"data_clean.csv")
```

```{r}
dflong <- df %>% 
  transmute(across(everything(), as.character)) %>% 
  pivot_longer(
    cols = -c(CKD),
    names_to = "colname",
    values_to = "val"
  )

write_csv(dflong,"data_long.csv")
```



# EDA

## Non-factor (Numerics)

### Age

```{r}
fig1a <- df %>% 
  group_by(Age,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = Age,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig1b <- df %>% 
  group_by(Age,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = Age,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig1a | fig1b) + 
  plot_annotation(
    title = "Figure 1: CKD dist. across Age by Count and Proportion"
  )
```

### BMI

```{r}
fig2a <- df %>% 
  mutate(BMI = round(BMI)) %>% 
  group_by(BMI,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = BMI,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig2b <- df %>% 
  mutate(BMI = round(BMI)) %>% 
  group_by(BMI,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = BMI,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig2a | fig2b) + 
  plot_annotation(
    title = "Figure 2: CKD dist. across BMI (rounded) by Count and Proportion"
  )
```

### Waist

```{r}
fig3a <- df %>% 
  mutate(Waist = round(Waist)) %>% 
  group_by(Waist,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = Waist,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig3b <- df %>% 
  mutate(Waist = round(Waist)) %>% 
  group_by(Waist,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = Waist,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig3a | fig3b) + 
  plot_annotation(
    title = "Figure 3: CKD dist. across Waist (rounded) by Count and Proportion"
  )
```

### SBP

```{r}
fig4a <- df %>% 
  mutate(SBP = round(SBP)) %>% 
  group_by(SBP,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = SBP,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig4b <- df %>% 
  mutate(SBP = round(SBP)) %>% 
  group_by(SBP,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = SBP,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig4a | fig4b) + 
  plot_annotation(
    title = "Figure 4: CKD dist. across SBP (rounded) by Count and Proportion"
  )
```

### DBP

```{r}
fig5a <- df %>% 
  mutate(DBP = round(DBP)) %>% 
  group_by(DBP,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = DBP,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig5b <- df %>% 
  mutate(DBP = round(DBP)) %>% 
  group_by(DBP,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = DBP,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig5a | fig5b) + 
  plot_annotation(
    title = "Figure 5: CKD dist. across DBP (rounded) by Count and Proportion"
  )
```

### HDL

```{r}
fig6a <- df %>% 
  mutate(HDL = round(HDL)) %>% 
  group_by(HDL,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = HDL,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig6b <- df %>% 
  mutate(HDL = round(HDL)) %>% 
  group_by(HDL,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = HDL,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig6a | fig6b) + 
  plot_annotation(
    title = "Figure 6: CKD dist. across HDL (rounded) by Count and Proportion"
  )
```

### LDL 

```{r}
fig7a <- df %>% 
  mutate(LDL = round(LDL)) %>% 
  group_by(LDL,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = LDL,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col(position = "fill")+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Proportion",
       fill = "CKD",
       col = "CKD")+ 
  theme(legend.position = "none")

fig7b <- df %>% 
  mutate(LDL = round(LDL)) %>% 
  group_by(LDL,CKD) %>% 
  summarise(n=n()) %>% 
  ggplot()+
  aes(x = LDL,
      y = n,
      fill = CKD %>% factor(),
      col = CKD %>% factor())+
  geom_col()+
  scale_fill_viridis_d() + scale_color_viridis_d()+
  labs(y = "Count",
       fill = "CKD",
       col = "CKD")

(fig7a | fig7b) + 
  plot_annotation(
    title = "Figure 7: CKD dist. across LDL (rounded) by Count and Proportion"
  )
```

## Factors

(Tableau crosstabs here)

# Modeling 1: Exploring Methods

## Splitting
```{r}
set.seed(8675309)

splits <- initial_split(df,
                        prop = .75,
                        strata = CKD)

d.train <- training(splits)
d.test  <- testing(splits)
```

## Preprocessing Steps

```{r}
recipe.d <- recipe(CKD~., data = d.train) %>% 
  step_relevel(
    Racegrp,
    ref_level = "white"
  ) %>% 
  step_relevel(
    CareSource,
    ref_level = "noplace"
  ) %>% 
  step_relevel(
    Activity,
    ref_level = "1"
  ) %>% 
  step_dummy(
    Racegrp,
    CareSource, 
    Activity
  ) %>% 
  step_zv(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors())
```

## Base GLM

### Training
```{r}
model.logreg <- 
  logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

wf.logreg <-
  workflow() %>% 
  add_model(model.logreg) %>% 
  add_recipe(recipe.d)

fit.logreg <- wf.logreg %>% fit(d.train)
```

### Coefficient-based VarImp

```{r}
tidy(fit.logreg) %>% 
  mutate(odds_ratio = round(exp(estimate),2)) %>% 
  mutate(sig        = case_when(p.value <= .001 ~ "***",
                                p.value <= .01 & p.value > 0.001 ~ "**",
                                p.value <= .05 & p.value > 0.01 ~ "*")) %>% 
  select(term, odds_ratio, p.value, sig) %>% 
  filter(term != "(Intercept)") %>% 
  arrange(p.value)
```

### Performance

```{r}
fit.logreg.test <- wf.logreg %>% last_fit(splits)

preds <- fit.logreg.test %>% 
  collect_predictions()
```

#### Red Herring: Unadjusted for imbalance

```{r}

metrics1f <- preds %>% pr_auc(truth = CKD,
                 estimate = .pred_0,
                 event_level = "first") %>%
  rbind(preds %>%
          f_meas(
            truth = CKD,
            estimate = .pred_class,
            event_level = "first"
          )) %>% 
  select(-.estimator)


# PR Curve
a <- preds %>% 
  pr_curve(CKD,.pred_0,
           event_level = "first") %>% 
  autoplot()

# ROC Curve
b <- preds %>% 
  roc_curve(
    truth = CKD,
    estimate = .pred_0,
    event_level = "first"
  ) %>% 
  autoplot()

(a | b) + 
  plot_annotation(
    title = paste("AUC of PR Curve: ",metrics1f[1,2],"\n",
                  "F1 Measure: ",metrics1f[2,2],sep = "")
  )
```

#### Truth: Adjusted for imbalance

```{r}
# PR-AUC + bACC
metrics1t <- preds %>% pr_auc(truth = CKD,
                 estimate = .pred_1,
                 event_level = "second") %>%
  rbind(preds %>%
          bal_accuracy(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  rbind(preds %>%
          sens(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  select(-.estimator)

a <- preds %>% 
  pr_curve(CKD,.pred_1,
           event_level = "second") %>% 
  autoplot()

b <- preds %>% 
  roc_curve(
    truth = CKD,
    estimate = .pred_1,
    event_level = "second"
  ) %>% 
  autoplot()

results_naive <- (a | b) + 
  plot_annotation(
    title = paste("AUC of PR Curve: ",metrics1t[1,2],"\n",
                  "Balanced Accuracy: ",metrics1t[2,2],"\n",
                  "True Positive Rate: ",metrics1t[3,2],
                  sep = "")
  )
```

## Subsampling 1: Downsampling

### Downsampling 

```{r}
recipe.downsample <- recipe.d %>% 
  step_downsample(CKD)
```

### Training

```{r}
model.logreg <- 
  logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

wf.logreg.downsample <-
  workflow() %>% 
  add_model(model.logreg) %>% 
  add_recipe(recipe.downsample)

fit.logreg.downsample <- wf.logreg.downsample %>% fit(d.train)
```

### Coefficient-based VarImp

```{r}
tidy(fit.logreg.downsample) %>% 
  mutate(odds_ratio = round(exp(estimate),2)) %>% 
  mutate(sig        = case_when(p.value <= .001 ~ "***",
                                p.value <= .01 & p.value > 0.001 ~ "**",
                                p.value <= .05 & p.value > 0.01 ~ "*")) %>% 
  select(term, odds_ratio, p.value, sig) %>% 
  filter(term != "(Intercept)") %>% 
  arrange(p.value)
```
### Performance

```{r}
fit.logreg.test.downsample <- wf.logreg.downsample %>% last_fit(splits)

preds <- fit.logreg.test.downsample %>% 
  collect_predictions()

# PR-AUC + bACC
metrics1t <- preds %>% pr_auc(truth = CKD,
                 estimate = .pred_1,
                 event_level = "second") %>%
  rbind(preds %>%
          bal_accuracy(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  select(-.estimator)

a <- preds %>% 
  pr_curve(CKD,.pred_1,
           event_level = "second") %>% 
  autoplot()

b <- preds %>% 
  roc_curve(
    truth = CKD,
    estimate = .pred_1,
    event_level = "second"
  ) %>% 
  autoplot()

(a | b) + 
  plot_annotation(
    title = paste("AUC of PR Curve: ",metrics1t[1,2],"\n",
                  "Balanced Accuracy: ",metrics1t[2,2],sep = "")
  )

```

## Subsampling 2: Upsampling

### Upsampling 

```{r}
recipe.upsample <- recipe.d %>% 
  step_upsample(CKD)
```

### Training

```{r}
wf.logreg.upsample <-
  workflow() %>% 
  add_model(model.logreg) %>% 
  add_recipe(recipe.upsample)

fit.logreg.upsample <- wf.logreg.upsample %>% fit(d.train)
```

### Coefficient-based VarImp

```{r}
tidy(fit.logreg.upsample) %>% 
  mutate(odds_ratio = round(exp(estimate),2)) %>% 
  mutate(sig        = case_when(p.value <= .001 ~ "***",
                                p.value <= .01 & p.value > 0.001 ~ "**",
                                p.value <= .05 & p.value > 0.01 ~ "*")) %>% 
  select(term, odds_ratio, p.value, sig) %>% 
  filter(term != "(Intercept)") %>% 
  arrange(p.value)
```
### Performance

```{r}
fit.logreg.test.upsample <- wf.logreg.upsample %>% last_fit(splits)

preds <- fit.logreg.test.upsample %>% 
  collect_predictions()

# PR-AUC + bACC
metrics1t <- preds %>% pr_auc(truth = CKD,
                 estimate = .pred_1,
                 event_level = "second") %>%
  rbind(preds %>%
          bal_accuracy(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  select(-.estimator)

a <- preds %>% 
  pr_curve(CKD,.pred_1,
           event_level = "second") %>% 
  autoplot()

b <- preds %>% 
  roc_curve(
    truth = CKD,
    estimate = .pred_1,
    event_level = "second"
  ) %>% 
  autoplot()

(a | b) + 
  plot_annotation(
    title = paste("AUC of PR Curve: ",metrics1t[1,2],"\n",
                  "Balanced Accuracy: ",metrics1t[2,2],sep = "")
  )

```

## Subsampling 3: SMOTE

### SMOTE 

```{r}
recipe.smote <- recipe.d %>% 
  step_smote(CKD)
```

### Training

```{r}
wf.logreg.smote <-
  workflow() %>% 
  add_model(model.logreg) %>% 
  add_recipe(recipe.smote)

fit.logreg.smote <- wf.logreg.smote %>% fit(d.train)
```

### Coefficient-based VarImp

```{r}
tidy(fit.logreg.smote) %>% 
  mutate(odds_ratio = round(exp(estimate),2)) %>% 
  mutate(sig        = case_when(p.value <= .001 ~ "***",
                                p.value <= .01 & p.value > 0.001 ~ "**",
                                p.value <= .05 & p.value > 0.01 ~ "*")) %>% 
  select(term, odds_ratio, p.value, sig) %>% 
  filter(term != "(Intercept)") %>% 
  arrange(p.value)
```
### Performance

```{r}
fit.logreg.test.smote <- wf.logreg.smote %>% last_fit(splits)

preds <- fit.logreg.test.smote %>% 
  collect_predictions()

# PR-AUC + bACC
metrics1t <- preds %>% pr_auc(truth = CKD,
                 estimate = .pred_1,
                 event_level = "second") %>%
  rbind(preds %>%
          bal_accuracy(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  select(-.estimator)

a <- preds %>% 
  pr_curve(CKD,.pred_1,
           event_level = "second") %>% 
  autoplot()

b <- preds %>% 
  roc_curve(
    truth = CKD,
    estimate = .pred_1,
    event_level = "second"
  ) %>% 
  autoplot()

(a | b) + 
  plot_annotation(
    title = paste("AUC of PR Curve: ",metrics1t[1,2],"\n",
                  "Balanced Accuracy: ",metrics1t[2,2],sep = "")
  )

```

# Modeling 2: Putting it all together

## Splitting

```{r}
set.seed(8675309)

splits <- initial_split(df,
                        prop = .75,
                        strata = CKD)

d.other <- training(splits)
d.test  <- testing(splits)

val_set <- validation_split(d.other,
                            prop = .8,
                            strata = CKD)
```

## Tuning 

```{r}
recipe <- recipe(CKD ~ ., data = d.other) %>%
  step_relevel(Racegrp,
               ref_level = "white") %>%
  step_relevel(CareSource,
               ref_level = "noplace") %>%
  step_relevel(Activity,
               ref_level = "1") %>%
  step_dummy(Racegrp,
             CareSource,
             Activity) %>%
  step_zv(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_smote(CKD)
  
model <- 
  logistic_reg(
    penalty = tune(),
    mixture = tune()
  ) %>% 
  set_engine("glmnet")

wf <- workflow() %>% 
  add_model(model) %>% 
  add_recipe(recipe)

grid <- expand.grid(
  penalty = c(0,.001,.01,.025,.05,.1,.2,.25,.5,1),
  mixture = c(0,0.25,0.5,0.75,1)
)

set.seed(8675309)
tune <- wf %>% 
  tune_grid(val_set,
            grid = grid,
            control = control_grid(save_pred = T),
            metrics = metric_set(bal_accuracy))

best <- tune %>% 
  select_best(metric = "bal_accuracy")

tune %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty,
             y = mean,
             col = mixture %>% factor())) +
  geom_point()+
  geom_line()+
  labs(
    y = "bACC",
    col = "Reg. Type"
  )+
  scale_color_viridis_d()+
  theme_minimal()
```
## Fitting 

```{r}
model.final <- 
  logistic_reg(penalty = best$penalty,
               mixture = best$mixture) %>% 
  set_engine("glmnet")

wf.final <-
  workflow() %>% 
  add_model(model.final) %>% 
  add_recipe(recipe)

set.seed(8675309)
fit.final <- wf.final %>% 
  last_fit(splits)
```

```{r}
fit <- wf.final %>% fit(d.other)

tidy(fit) %>% 
  select(-penalty) %>% 
  filter(term != "(Intercept)") %>% 
  arrange(desc(abs(estimate)))
```


## Performance

```{r}
preds <- fit.final %>% 
  collect_predictions()

# PR-AUC + bACC
metrics1t <- preds %>% pr_auc(truth = CKD,
                 estimate = .pred_1,
                 event_level = "second") %>%
  rbind(preds %>%
          bal_accuracy(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  rbind(preds %>%
          sens(
            truth = CKD,
            estimate = .pred_class,
            event_level = "second"
          )) %>% 
  select(-.estimator)

a <- preds %>% 
  pr_curve(CKD,.pred_1,
           event_level = "second") %>% 
  autoplot()

b <- preds %>% 
  roc_curve(
    truth = CKD,
    estimate = .pred_1,
    event_level = "second"
  ) %>% 
  autoplot()

results_final <- (a | b) + 
  plot_annotation(
    title = paste("AUC of PR Curve: ",metrics1t[1,2],"\n",
                  "Balanced Accuracy: ",metrics1t[2,2],"\n",
                  "True Positive Rate: ",metrics1t[3,2],
                  sep = "")
  )
```
```{r}
results_naive
results_final
```

